A Graph-based Framework for Coverage Analysis in Autonomous Driving

### 2025-09-28 article/article.tex
small additions to article
small part of paper writing

>>>>

\section{Implementation of Graph Embeddings for Traffic Scene Analysis}

This section describes the implementation and application of graph embeddings to traffic scene graphs, as demonstrated in the \texttt{graph\_embeddings.ipynb} notebook. The implementation follows a comprehensive approach to learning graph representations through self-supervised contrastive learning.

\subsection{Graph Neural Network Architecture}

The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings. The model consists of multiple GINE convolutional layers, each containing:
\begin{itemize}
    \item Sequential multi-layer perceptrons (MLPs) with batch normalization and ReLU activations
    \item Edge-aware message passing incorporating both node and edge features
    \item Dropout regularization to prevent overfitting
\end{itemize}

The model employs a combination of three graph pooling strategies - mean, max, and sum pooling - to create a comprehensive graph-level representation. A projection head enables contrastive learning through InfoNCE loss, while an optional classification head supports supervised learning tasks.

\subsection{Data Processing and Augmentation}

The \texttt{GraphDataset} class handles the conversion from NetworkX graphs to PyTorch Geometric format through the \texttt{networkx\_to\_pyg} function. This conversion includes:
\begin{itemize}
    \item One-hot encoding of categorical node features (actor types: VEHICLE, PEDESTRIAN, CYCLIST, MOTORCYCLE)
    \item One-hot encoding of edge types (neighbor\_vehicle, opposite\_vehicle, same\_lane, adjacent\_lane, following, intersection)
    \item Preservation of continuous features such as longitudinal speed and path length
\end{itemize}

Data augmentation is implemented through the \texttt{augment\_graph} function, which adds Gaussian noise to continuous features to create augmented versions for contrastive learning.

\subsection{Training Methodology}

The training process utilizes self-supervised contrastive learning with the following characteristics:
\begin{itemize}
    \item InfoNCE contrastive loss implemented in the \texttt{contrastive\_loss} function
    \item Adaptive learning rate schedule with exponential decay (starting at 0.02, decaying by factor 0.75)
    \item Multi-epoch training with 15 epochs per learning rate stage
    \item Data from both CARLA simulator (4,442 graphs) and Argoverse 2.0 dataset (3,588 graphs)
\end{itemize}

\subsection{Embedding Analysis and Visualization}

The trained model generates 256-dimensional embeddings for each traffic scene graph. Analysis includes:
\begin{itemize}
    \item Dimensionality reduction using Principal Component Analysis (PCA) and t-distributed Stochastic Neighbor Embedding (t-SNE)
    \item Visualization of embedding space distinguishing between CARLA and Argoverse data sources
    \item Similarity-based scene retrieval using squared Euclidean distance in embedding space
    \item Integration with visualization functions (\texttt{plot\_lane\_map\_advanced}, \texttt{add\_actors\_to\_map}, \texttt{add\_actor\_edges\_to\_map}) to display similar traffic scenes
\end{itemize}

The implementation demonstrates successful learning of traffic scene representations, with the embedding space capturing meaningful similarities between scenes from different data sources while maintaining distinction between simulation and real-world data.
